RE: Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2023-05-17 rmd/manuscript.Rmd
edit

In the chain motif, productivity greatly increased the overall FCL (**Figure S5**), thereby shifting the upper bound of FCL (**Figures 3 and 4**). In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 3, 4, and S5**). This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 3 and 4**). State transitions associated with increased productivity depended on the motif structure. In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S8**). The introduction of omnivory allowed the transition between "B + P" and "B + C + P". With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S8**). However, this transition was marginal in the strong omnivory motif. Instead, productivity significantly increased the proportion of the "No species" state (**Figure S8**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures S5 and S6**).

Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S5**). State transitions occurred through an additive (chain) or insertion mechanism (omnivory). In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S9**). In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S9**). However, this direct effect was weak in comparison to other factors with no qualitative alteration to the relationships between FCL and ecosystem geometry (**Figures 3 and 4**).

@@

knitr::kable(df_value %>%
               dplyr::select(-"One-patch") %>%
               rename(Value = "Main"),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Parameter values used in the main simulation") %>%
  kableExtra::kable_styling(latex_options = 'HOLD_position')
```

>>>>

In the chain motif, productivity greatly increased the overall FCL (**Figure S5**), thereby shifting the upper bound of FCL (**Figures 3 and 4**). In contrast, increased productivity decreased FCL with weak and strong omnivory (**Figures 3, 4, and S5**). This negative effect of increased productivity was apparent in smaller and/or simpler ecosystems, steepening the slopes between FCL and ecosystem geometry (**Figures 3 and 4**). State transitions associated with increased productivity depended on the motif structure. In the chain motif, the transition from "B + C" to "B + C + P" was dominant regardless of disturbance frequency (**Figure S6**). The introduction of omnivory allowed the transition between "B + P" and "B + C + P". With weak omnivory, the transition of "B + C + P" $\rightarrow$ "B + P" was prevalent and shortened FCL across disturbance regimes (**Figure S6**). However, this transition was marginal in the strong omnivory motif. Instead, productivity significantly increased the proportion of the "No species" state (**Figure S6**) in disturbed scenarios, especially in small and/or simple ecosystems (**Figures 5 and 6**).

Dispersal distance had a positive influence on FCL regardless of simulation scenarios (**Figure S5**). State transitions occurred through an additive (chain) or insertion mechanism (omnivory). In the chain motif, the transitions were either "B" $\rightarrow$ "B + C" or "B + C" $\rightarrow$ "B + C+ P", depending on disturbance frequency (**Figure S7**). In the omnivory motifs, the predominant transition was from "B + P" to "B + C + P" (**Figure S7**). However, this direct effect was weak in comparison to other factors with no qualitative alteration to the relationships between FCL and ecosystem geometry (**Figures 3 and 4**).

@@

knitr::kable(df_value %>%
               dplyr::select(-"One-patch") %>%
               rename(Value = "Main"),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Parameter values used in the main simulation")
```

### 2023-05-17 rmd/reply2.Rmd
edit

This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" Thank you for allowing us to revise our manuscript. The comments were very helpful in improving our manuscript. When revising the manuscript, we have put a particular emphasis on the following points:

XXX

>>>>

This is a re-submission of a manuscript: ECS22-0646 "*Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems.*" We appreciate the time you have taken for our manuscript. When revising the manuscript, we have put a particular emphasis on the following points:

1. Reorganized the results section to better focus on ecosystem size/complexity effects
2. Improved figure captions

### 2023-05-17 rmd/tables_figures.Rmd
edit rmd
edit

**Figure 1** (A) Example of simulated branching networks. Node color is reflective of the disturbance value calculated in network generation. (B) Density distributions of disturbance intensity (i.e., proportional reduction of a population). Three networks of 50 patches were produced with varied branching probabilities, which were shown in different colors. The vertical line indicates the mean disturbance intensity at headwaters ($\mu_m = 0.8$). (C) Trophic interactions in a three-species community. $B$ refers to a basal species whose population level is controlled by a density-dependent function. The consumer species (or intraguild prey), $C$, consumes $B$ as a resource. The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$. The structure and strength of these trophic interactions determined by functional responses. On the far left, a simple linear food chain exists when the predator does not consume the basal species. Moving to the right, basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates. (D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios. The contribution changes with the relative density of $B$ to $C$ ($C=50$ in this example), and the rate of change is determined by attack rate and handling time.

@@

```{r figure2, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_main_n_patch.pdf"))
```

**Figure 2** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$. Lines indicate loess curves, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities.

@@

```{r figure3, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_main_p_branch.pdf"))
```

**Figure 3** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$. Lines indicate loess curves, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities.

>>>>

**Figure 1** (A) Example of simulated branching networks. Node color reflects the disturbance value calculated during network generation. (B) Density distributions of disturbance intensity (i.e., proportional reduction of a population). Three networks of 50 patches were produced with varied branching probabilities, shown in different colors. The vertical line indicates the mean disturbance intensity at the headwaters ($\mu_m = 0.8$). (C) Schematic diagram of trophic interactions in a three-species community. $B$ refers to a basal species whose population level is controlled by a density-dependent function. The consumer species (or intraguild prey), $C$, consumes $B$ as a resource. The predator (or intraguild predator), $P$, is able to consume both $B$ and $C$. The structure and strength of these trophic interactions are determined by functional responses. On the far left, a simple linear food chain exists when the predator does not consume the basal species. Moving to the right, the basal species' biomass is converted to predator biomass at higher efficiencies with greater attack rates. Arrow colors schematically indicate the relative reliance of $P$ on $B$ and $C$, with black indicating trophic specialization. (D) Proportional contribution of basal species $B$ to predator $P$ under different omnivory scenarios. The contribution changes with the relative density of $B$ to $C$ ($C = 50$ in this example), and the rate of change is determined by attack rate and handling time.

@@

```{r figure2, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_tree.pdf"))
```

**Figure 2** Summary of our simulation results. Black text in each node indicates the potential factors that may influence the association between food chain length and ecosystem geometry, and black text on arrows indicates simulated scenarios for a given factor. Blue text denotes the resultant association between food chain length and ecosystem geometry. Panel (A) indicates the effect of ecosystem size. Panel (B) indicates the effect of ecosystem complexity.

@@

```{r figure3, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_main_n_patch.pdf"))
```

**Figure 3** Relationships between food chain length and ecosystem size with high disturbance intensity $\mu_m = 0.8$. Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities. In this figure, ecosystem complexity values for individual simulation replicates were drawn randomly from a uniform distribution $P_b \sim \text{Unif}(0.01, 0.99)$.

```{r figure4, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_main_p_branch.pdf"))
```

**Figure 4** Relationships between food chain length and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$. Lines in each panel indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals. Rows distinguish food web motifs, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities. In this figure, ecosystem size values for individual simulation replicates were drawn randomly from a uniform distribution $N_p \sim \text{Unif}(10, 150)$.

```{r figure5, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_state_n_patch.pdf"))
```

**Figure 5** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem size with high disturbance intensity $\mu_m = 0.8$. The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate. Lines indicate loess curves fitted to $500$ simulation replicates, and shades are associated 95% confidence intervals. Panels indicate different food web motifs. In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities.

```{r figure6, fig.width=12, fig.height=5.5}
knitr::include_graphics(here::here("figure/figure_state_p_branch.pdf"))
```

**Figure 6** Relationships between the proportion of each patch state $\bar{\psi}$ (colors indicate different patch states) and ecosystem complexity with high disturbance intensity $\mu_m = 0.8$. The proportion of each patch state $\bar{\psi}$ is averaged over time for each simulation replicate. Lines indicate loess curves fitted to 500 simulation replicates, and shades are associated with 95% confidence intervals. Panels indicate different food web motifs. In each panel, rows distinguish disturbance probability $p_m$, and columns indicate different dispersal and productivity scenarios. Different line colors indicate different disturbance probabilities.
